maASlin2

### 2019-12-09 scripts/27_maASlin2.Rmd
updte

```{r}
# now change the columns to different type: dbl or factor
pheno_final  <- pheno %>%  
  # change the ferritin to dbl
  mutate(ferritin = as.numeric(ferritin),
         Key__Patient_Disease_Status_at_CAR = as.character(Key__Patient_Disease_Status_at_CAR)) %>% 
  # change all the pheno vars to factors
  mutate_if(is.character, as.factor) %>% 
  mutate(MRN = as.character(MRN),
         sampleid = as.character(sampleid)) %>% 
  column_to_rownames('sampleid')

@@

fit_data <- Maaslin2(
    pw, pheno_final, '../data/maASlin2/pw', normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = T,
    fixed_effects = c('Toxicity', 
                      'Cytokine_Release_Syndrome',
                      'ICANS',
                      'CR',
                      'ORR', 
                      'Disease_Class',
                      'Key__Patient_Disease_Status_at_CAR',
                      'Antibiotics_4_weeks_prior_to_CAR_T_cells', 
                      'ferritin',
                      'crp',
                      'IL_1_beta',
                      'IL_10',
                      'IL_6'),
    # random_effects = c('MRN',
    #                    'sex',
    #                    'age'),
    standardize = FALSE)

>>>>

```{r}
# now change the columns to different type: dbl or factor
pheno_final  <- pheno %>%  
  # change the ferritin to dbl
  mutate(ferritin = as.numeric(ferritin),
         Key__Patient_Disease_Status_at_CAR = as.character(Key__Patient_Disease_Status_at_CAR)) %>% 
  # change all the pheno vars to factors
  mutate_if(is.character, as.factor) %>% 
  mutate(MRN = as.character(MRN),
         sampleid = as.character(sampleid)) %>% 
  mutate(IL_6 = log(IL_6),
         IL_10 = log(IL_10)) %>% 
  column_to_rownames('sampleid')

@@

# filter on the above matrix to include pathways existing in at least 20% of the samples (>=4)
samp_num <- nrow(pw)

ok_samps <- colSums(pw != 0)

# to see how many samps pass the number of zeros threshold
pers <- seq(0.1, 0.99, 0.1)

pers %>% 
  set_names(pers) %>% 
  map_dfc(~ pw[,names(ok_samps[ok_samps >= ceiling(samp_num * .)])] %>% 
            ncol) %>% 
  gather( key = 'per', value = 'num')

# temporarily decides to choose 
per_thre <- 0.5

pw_fil <- pw[,names(ok_samps[ok_samps >= ceiling(samp_num * per_thre)])]

```{r}

fit_data <- Maaslin2(
    pw_fil, pheno_final, '../data/maASlin2/pw', normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = T,
    fixed_effects = c('Toxicity', 
                      'Cytokine_Release_Syndrome',
                      'ICANS',
                      'CR',
                      'ORR', 
                      'Disease_Class',
                      'Key__Patient_Disease_Status_at_CAR',
                      'Antibiotics_4_weeks_prior_to_CAR_T_cells', 
                      'ferritin',
                      'crp',
                      'IL_1_beta',
                      'IL_10',
                      'IL_6'),
    random_effects = c('sex',
                       'age'),
    standardize = FALSE)
